Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-10-06 Final_Manuscript.tex
edits, almost there

Barbara L Fredrickson and Roberts (1997) Objectification Theory,
suggests that, in addition to being steeped in a culture that
objectifies women, women are objectified in actual interpersonal
encounters. The negative effects of this interpersonal objectification
for women are theorized be the strongest when it is a man doing the
objectifying (Calogero, 2004; Gay \& Castano, 2010; Gervais, Vescio, \&
Allen, 2011), but what are the consequences when a woman objectifies
another woman in an interaction? Psychological researchers studying the
sexual objectification of women, have in recent years explored the
interpersonal process of the objectification, finding evidence for
\emph{self}-objectification in the target as a proximal consequence of
being objectified (Garcia et al., 2016; Riemer, Sáez, Brock, \& Gervais,
2020; Strelan \& Pagoudis, 2018). In a recent review, Gervais et al.
(2020) organized the extant literature on interpersonal
self-objectification and proposed a theoretical model called the Social
Interaction Model of Objectification (SIMO). The SIMO is useful for
studying mixed-gender interactions, but what about interactions among
women? Although there is evidence that women objectify other women
(Loughnan et al., 2015a; Puvia \& Vaes, 2013), studies investigating the
process of interpersonal objectification in interactions between women
is scarce.

Is objectification by a female interaction partner related to an
increases in self-objectification for the woman being objectified?
Further, does objectification by a woman have the same downstream
negative consequences for women as being objectified by a male
interaction partner? The current study begins to address these questions
by investigating interpersonal objectification among interacting pairs
where both partners identify as women.

@@

Self-objectification is a psychological process that provides that
translates the experiences of interpersonal objectification {[}Loughnan
et al. (2015a); Gervais et al. (2020) into negative mental health
outcomes (e.g., interaction inauthenticity, anxiety, self-esteem, and
cognitive performance) (Calogero, Tantleff-Dunn, \& Thompson, 2011;
Moradi \& Huang, 2008). Calogero et al. (2011) proposes that
self-objectification can be conceptualized as a learned trait, or
\emph{trait} self-objectification (TSO), whereby one abopts a habitual
third-person perspective on one's own appearance. Furthermore,
self-objectification can also be elicited momentarily, for example, when
viewing sexualized images in movies and magazines (Morry \& Staska,
2001), when trying on sexualizing clothing (Barbara L Fredrickson,
Roberts, Noll, Quinn, \& Twenge, 1998), or during video chat when you
believe your interaction partner can only see your body (not face;
Saguy, Quinn, F Dovidio, \& Pratto, 2010). This momentary
self-objectification is referred to as \emph{state} self-objectification
(SSO; Bernard, Legrand, \& Klein, 2018; Calogero et al., 2011; Moradi \&
Huang, 2008) and is characterized by feeling like a body rather than a
full self within a particular moment, instance, or context.

There is evidence that being objectified by another person during an
interaction elicits SSO. For example, Loughnan, Baldissarri, Spaccatini,
and Elder (2017) found that for women, imagining a time when they were
objectified by another person cause reductions in human traits
attributed to the self, and further, the gender of the person doing the
objectification was not a moderating factor. In addition, in a
face-to-face mixed-gender dyadic study, Garcia et al. (2016) found that
men's reported objectification of their female interaction partner was
associated with increased self-objectification as reported by their
female partner. Further, this negative effect of the men's
objectification on women's SSO was strongest for women higher in TSO.
The current study uses this same dyadic paradigm to study same-gender,
woman-woman interacting pairs.

@@

Past research investigating women objectifying other women has focused
on the psychological conditions that led women to objectify women.
Harsey and Zurbriggen (2020) found that self-objectification was related
to the objectification of women to a similar degree for male and female
participants. Along the same lines, Strelan and Hargreaves (2005) found
that the more women self-objectify, the more they objectify other women.
Could the process work in the other direction as well, that is, could
objectification by a woman produce self-objectification, in a similar
manner to objectification by a man? Hill and Fischer (2008) assessed
women's experiences of objectification from men independently from
women's experiences of objectification from other women. They found that
women may be socialized not only to see themselves as objects, but
perhaps to see other women as objects as well. This process, however,
was not comfirmed to occur during an actual interpersonal interaction.
Loughnan et al. (2015a) found that women objectify other women to a
greater extent than they objectify men. Women have also been found to
objectify (dehumanized) sexualized targets presented as images (Puvia \&
Vaes, 2013), but very little is known about interpersonal sexual
objectification among women and what the proximal effects of this
objectification might be. We do know that objectification has more
adverse consequences for women than men (Gervais et al., 2011; Moradi \&
Huang, 2008; Saguy et al., 2010), however, if women objectify other
women, perhaps even to a stronger degree than they objectify men, this
woman to woman objectification might have detrimental outcomes.

@@

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), often times feel \enquote{looked at}
within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and are likely to internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Interpersonal objectification has been found to have negative
consequences for women. Gervais et al. (2011) found that an objectifying
gaze by a male interaction partner (confederate) was associated with
lower math performance for women, than being objectified by a female
interaction partner. Perhaps the most adverse negative consequence of
interpersonal objectificaiton is that being objectified socializes girls
and women to routinely treat themselves as objects to be looked at and
evaluated (Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972; Barbara L
Fredrickson et al., 1998). Gervais et al. (2020) recently reviewed the
research on interpersonal objectification and organized our current
theoretical understanding of this process in the Social Interaction
Model of Objectification (SIMO). What is clear from Gervais et al.
(2020)'s review is that we know little about women-on-women
interpersonal objectification. Could there also be negative consequences
for women of interpersonal objectification by women?

Researchers have also studied interpersonal objectification and
self-objectification in romantic relationships (Strelan \& Pagoudis,
2018), in addition to interactions with strangers. Among heterosexual
male and female college students, self-objectification has been found to
be positively associated with the extent to which they objectify their
romantic partners (Zurbriggen, Ramsey, \& Jaworski, 2011). Evidence is
mixed, suggesting that women can sometimes have positive outcomes from
appearance valuation within the context of an established romantic
relationship. There is also evidence that they experience less SSO after
appearance comments within romantic relationships (Meltzer, 2020), but
less is known about first dates. An important caveat to this literature
is that it is only about heterosexual relationships. A recent study has
investigated the effect of self-objectification on the reduction of
relationship building skills in general (including same-sex friendships)
(Terán, Jiao, \& Aubrey, 2020). Both romantic relationships and stranger
encounters have been studied in the context of objectification, but
neither of these literatures have investigated specifically same-gender
relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
for women would suggest no (Calogero, 2004; Gay \& Castano, 2010;
Gervais et al., 2011; Saguy et al., 2010; Yilmaz \& Bozo, 2019), but,
there is evidence that women do objectify other women (Harsey \&
Zurbriggen, 2020; Loughnan et al., 2015a; Puvia \& Vaes, 2013) and to
would be helpful to understand the consequencs of this intragroup
objectification for women. When women objectify other women, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

There is some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán et al., 2020). This
link has been justified by the literature on stigmatized-stigmatizer
interactions (M. R. Hebl \& Dovidio, 2005; Richeson \& Shelton, 2003;
Shelton, Richeson, \& Salvatore, 2005), viewing the experience of being
objectified in an interaction as an identity threat situation
(Miles-McLean et al., 2015; Nadal \& Haynes, 2012). Further, empirical
evidence reveals that objectification manifests through inauthenticity
in romantic relationships (Brunell et al., 2010), adverse attitudes in
regard to career aspirations, and a decrease in concentration and
impairment in female cognitive performance (Kahalon, Shnabel, \& Becker,
2018; D. M. Quinn, Chaudoir, \& Kallen, 2011). When a woman is
objectified by a man, and subsequently experiences self-objectification,
the intergroup nature of this encounter might trigger identity threat.
Thus, perhaps there are fewer negative consequences when a woman is
objectified by another woman, specifically, woman-woman interpersonal
objectification processes might diverge from mix-gender interpersonal
objectification processes precisely because they are not intergroup
interactions (at least with respect to gender identity). A woman
objectified by another woman may not be having the same negative
consequences that cascade from situations that trigger group-based
identity threat (Deaux \& Major, 1987; Dovidio, Hebl, Richeson, \&
Shelton, 2006; M. R. Hebl \& Dovidio, 2005), but feeling like a body,
rather than a full human, in any interaction (intergroup or ingroup) may
be enough to reduce women's feelings of authenticity and social
competence, regardless of the gender of the objectifyer (Terán et al.,
2020; Tolman, Impett, Tracy, \& Michael, 2006).

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected that TSO
would moderate this relationship, amplifying the positive assocaition
between being objectified and SSO for women higher in TSO. Here we would
like to note that Puvia and Vaes (2013) would alternatively predict that
women's tendency to self-objectify (TSO) leads them to objectify other
women (more precisely, to dehumanize a sexualized woman). In addition,
Strelan and Hargreaves (2005) and Harsey and Zurbriggen (2020) would
predict the TSO to other-objectification link, further theorizing that
this relationships is mediated by the woman's own state
self-objectification (SSO). This limited, but extant, set of literature
on woman-woman objectification points to a possible alternative model
from the model hypothesized in the current study. Where appropriate, we
report results considering the causal directions implied by this
alternative mediation model.

We also tested whether SSO would, in turn, lead to feelings of
inauthenticity. The investigation of the SSO - inauthenticity connection
is considered exploratory given the lack of support for this connection
from the literature on identity threat in intragroup interactions
(Rollero, 2016) but support for this connection in the objectification
literature (Garcia et al., 2016; Terán et al., 2020). Regardless of
whether there was an association between SSO and inauthenticity, we
hypothesized that feelings of inauthencity would be associated with
reduced feelings of agency in romantic relationships, reduced career
aspiration, and reduced cognitive performance. In summary, we expected
to find a positive relationship between other-objectification by one's
partner and state self-objectification. We also expected to find a
negative relationship between self-state objectification and interaction
authenticity, and that interaction authenticity will be positively
related to cognitive performance, relationship agency, and career
aspirations.

@@

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Westurn culture and beyond, women need to
look attractive to obtain and maintain successful relationship, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners (i.e., their sexual attractiveness; Puvia \& Vaes,
2013)---indeed, women may be unfortunately quite used to thinking about
their own \enquote{datability,} and we suspect this this habitual
thought pattern will translate to their thoughts about other women.

@@

Thirty-two previously unacquainted dyads of self-identifying women
participated in this study. Data from two demographically similar higher
education institutions in the Northeast United States were combined to
create the final analysis sample (\emph{N =} 64) used in this study. In
the measures section that follows we refer to them as Sample 1 and
Sample 2. Sample 1 (\emph{N =} 24) is from a co-ed liberal arts college
and Sample 2 (\emph{N =} 40) is from a women's liberal arts college.
More specifically, twelve of the pairs were students at a co-ed liberal
arts college, while the remaining twenty pairs attended a women's
liberal arts college.\footnote{Initially, data was collected from both
  same-gender and mixed-gender dyads at both institutions. Sample 1
  originally consisted of 22 pairs, 12 men and 32 women. In Sample 2
  there were 23 pairs made up of 43 women and one man, as well as two
  participants who did not identify with either gender category. To
  investigate only same-gender pairs of women, we limited participant
  data to women in both samples.} All results presented below are from
models including sample as a control variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were cross-racial minority group pairs. 64.06\%
of the sample identified as heterosexual, and 25\% identified as gay,
lesbian or bisexual. See the analysis of differences between
heteroseuxal and non-heterosexual women in levels of
other-objectification of their female partners in the measures section
below.

@@

The following measures were collected in the order they are presented
following the interaction. Correlations and descriptive statistics
appear in Table~\ref{tab:corrtable}.

@@

Trigrams from the Remote Associates Task (McFarlin \& Blascovich, 1984)
were utilized to assess cognitive performance after the interaction. Ten
items were selected and presented to participants. For example, the
correct answer for the trigram \enquote{Quack: Pond: Waddle} would be
\enquote{Duck}. Participants are limited to 30 seconds. For every
correct answer, 1 point is given. The mean score was 5.03 (SD = 2.29).
Cognitive performance was measured first in order to measure potential
immediate detriments to performance (Garcia et al., 2016).

@@

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts in relation to multiple characteristics of their
partner (Garcia et al., 2016). Questions included aspects of their
partner's internal traits such as personality, friends, family, and
extracurricular interests, as well as external traits such as body,
appearance, clothing, and body parts. All questions were rated on a
scale from 1 (not at all) to 7 (constantly). Objectification was
measured by taking the difference between the average frequency of
thought about their partner's external traits (\(\alpha\) = 0.79 for
Sample 1, \(\alpha\) = 0.79 for Sample 2) and frequency of thought about
their partner's internal traits (\(\alpha\) = 0.79 for Sample 1,
\(\alpha\) = 0.76 for Sample 2). A positive score in this scale would
indicate that the participant thought about their partner's external
traits more than the partner's internal traits, and a negative score
would indicate the opposite.

@@

To assess the magnitude to which individuals felt comfortable in the
interaction and perceived the interaction to be authentic, we asked
participants to rate the extent to which they felt comfortable, happy,
friendly, warm, easygoing, sincere, and authentic on a scale ranging
from 1 (not at all) to 7 (very much), much alike (Garcia et al., 2016).
Participants were additionally asked to rate their interaction partner's
authenticity as well as their own: \enquote{`Do you think your partner
was authentic during your interaction?}' and \enquote{`Were you
authentic during your interaction?}' These questions were ranked on a
scale from 1 (not at all) to 7 (very much). These were combined to form
the authenticity scale (\(\alpha\) = 0.91 for Sample 1, \(\alpha\) =
0.91 for Sample 2).

@@

A scale was used from Garcia et al. (2016) to assess how much agency an
individual believes they would possess in future romantic relationships.
Participants were asked how likely it was that they would do the
following: \enquote{`ask someone out on a date,}' \enquote{`open the
door for your date,}' \enquote{`pay for a date,}' \enquote{`ask your
boyfriend/girlfriend to marry you,}' \enquote{`initiate sex with your
girlfriend/boyfriend,}' \enquote{`initiate condom use during sex,}'
\enquote{`surprise your boyfriend/ girlfriend with a gift,}' and
\enquote{`ask your girlfriend/boyfriend to move with you to a new
place.}' Responses were measured on a scale ranging from 1 (not at all
likely) to 7 (extremely likely). The scale originally had 9 items, but
the 9th item had low correlations with the remaining items, ranging from
.02 to .30 for the first sample, and .04 to .30 for the second sample.
The item was intended to be reverse coded, but correlations were still
low enough to make the scale unreliable. Therefore, the ninth item was
removed. As a result, the scale had moderately high reliability for both
samples (\(\alpha\) = 0.72 for Sample 1, (\(\alpha\) = 0.74 for Sample
2).

@@

We used R (Version 3.5.2; R Core Team, 2017) and the R-packages
\emph{tidyverse} (Version 1.2.1; Wickham, 2017), \emph{psych} (Version
1.8.4; Revelle, 2017), \emph{nlme} (Version 3.1.137; Pinheiro, Bates,
DebRoy, Sarkar, \& R Core Team, 2017) for our analyses. Further, we used
the R-packages \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
\emph{apaTables} (Version 2.0.5; Stanley, 2018), with \emph{knitr}
(Version 1.25; Xie, 2015) to create a fully reproducible manuscript in
\emph{rmarkdown} (Version 1.15; Xie, Allaire, \& Grolemund, 2018).

>>>>

Fredrickson and Roberts (1997)'s Objectification Theory, suggests that,
in addition to being steeped in a culture that objectifies women, women
are objectified in actual interpersonal encounters. The negative effects
of this interpersonal objectification for women have been theorized to
be the strongest under the \enquote{male gaze,} that is, when it is a
percieved or actual man doing the objectifying (Calogero, 2004; Gay \&
Castano, 2010; Gervais, Vescio, \& Allen, 2011), but what are the
consequences when a woman objectifies another woman in an interaction?
Psychological researchers studying the sexual objectification of women
have in recent years explored the interpersonal process of the
objectification, finding evidence for \emph{self}-objectification in the
target as a proximal consequence of being objectified (Garcia et al.,
2016; Riemer, Sáez, Brock, \& Gervais, 2020; Strelan \& Pagoudis, 2018).
In a recent review, Gervais et al. (2020) organized the extant
literature on interpersonal self-objectification and proposed a
theoretical model called the Social Interaction Model of Objectification
(SIMO). The SIMO is useful for studying mixed-gender interactions, but
can it be extended to interactions among women? Although there is
evidence that women objectify other women (Loughnan et al., 2015a; Puvia
\& Vaes, 2013), studies investigating the process of interpersonal
objectification in interactions between women is scarce.

The current study uses a face-to-face interaction methodology to being
to answer a series of research questions about interpersonal
objectification among women: Is objectification by a female interaction
partner related to an increase in self-objectification for the woman
being objectified? Does objectification by a woman have the same
downstream negative consequences for women as being objectified by a
male interaction partner? The current study addresses these questions by
investigating interpersonal objectification among interacting pairs
where both partners identify as women. We first review the literature on
self-objectification, then we review the evidence for women objectifying
other women before turning to the literature on interpersonal
objectification.

@@

Self-objectification is a psychological process that translates the
experiences of interpersonal objectification (Gervais et al., 2020;
Loughnan et al., 2015a) into negative mental health outcomes (e.g.,
interaction inauthenticity, anxiety, lower self-esteem, and poor
cognitive performance; Calogero, Tantleff-Dunn, \& Thompson, 2011;
Moradi \& Huang, 2008). Calogero et al. (2011) proposes that
self-objectification can be conceptualized as a learned trait, or
\emph{trait} self-objectification (TSO), whereby one abopts a habitual
third-person perspective on one's own appearance. Furthermore,
self-objectification can also be elicited momentarily, for example, when
viewing sexualized images in movies and magazines (Morry \& Staska,
2001), when trying on sexualizing clothing (Barbara L Fredrickson,
Roberts, Noll, Quinn, \& Twenge, 1998), or during video chat when you
believe your interaction partner can only see your body (not face;
Saguy, Quinn, F Dovidio, \& Pratto, 2010). This momentary
self-objectification is referred to as \emph{state} self-objectification
(SSO; Bernard, Legrand, \& Klein, 2018; Calogero et al., 2011; Moradi \&
Huang, 2008) and is characterized by feeling like a body rather than a
full self within a particular moment, instance, or context.

There is evidence that being objectified by another person during an
interaction elicits SSO. For example, Loughnan, Baldissarri, Spaccatini,
and Elder (2017) found that for women, imagining a time when they were
objectified by another person caused reductions in human traits
attributed to the self, and notably, the gender of the person doing the
objectification was not a moderating factor. In addition, in a
mixed-gender dyadic study of actual interactions, Garcia et al. (2016)
found that men's reported objectification of their female interaction
partner was associated with increased self-objectification (as reported
by their female partner). Further, this negative effect of the men's
objectification on women's SSO was strongest for women higher in TSO.
The current study uses this same dyadic paradigm to study same-gender,
woman-woman interacting pairs.

@@

Past research investigating women objectifying other women has focused
on the psychological conditions that led women to objectify women.
Harsey and Zurbriggen (2020) found that (trait) self-objectification was
related to the objectification of women to a similar degree for male and
female participants. Along the same lines, Strelan and Hargreaves (2005)
found that the more women self-objectify, the more they objectify other
women. Could the process work in the other direction as well, that is,
could objectification by a woman produce (state) self-objectification,
in a similar manner to objectification by a man? Hill and Fischer (2008)
assessed women's experiences of objectification from men independently
from women's experiences of objectification from other women. They found
that women may be socialized not only to see themselves as objects, but
perhaps to see other women as objects as well. This process, however,
was not comfirmed to occur during an actual interpersonal interaction.
Thus, there is evidence that women do objectify other women, indeed,
Loughnan et al. (2015a) found that women objectify other women to a
greater extent than they objectify men. Women have also been found to
objectify (dehumanized) sexualized targets presented as images (Puvia \&
Vaes, 2013), but very little is known about interpersonal sexual
objectification among women and what the effects of this objectification
might be. Past studies have foud that objectification has more adverse
consequences for women than men (Gervais et al., 2011; Moradi \& Huang,
2008; Saguy et al., 2010), however, we do not know much about the effect
of the gender of the \emph{objectifyer} on potential detrimental
outcomes.

@@

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), often times feel \enquote{looked at}
within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and are likely to internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). As mentiond above, interpersonal objectification has been found
to have negative consequences for women. Gervais et al. (2011) found
that an objectifying gaze by a male interaction partner (confederate)
was associated with lower math performance for women than being
objectified by a female interaction partner. Perhaps the most adverse
negative consequence of interpersonal objectificaiton is that being
objectified socializes girls and women to routinely treat themselves as
objects to be looked at and evaluated (Bartky, 1990; Berger, Cohen, \&
Zelditch Jr, 1972; Barbara L Fredrickson et al., 1998). Gervais et al.
(2020) recently reviewed the research on interpersonal objectification
and organized our current theoretical understanding of this process in
the Social Interaction Model of Objectification (SIMO). What is clear
from Gervais et al. (2020)'s review is that we know little about
women-on-women interpersonal objectification. Could there also be
negative consequences for women after being objectified by a woman?

Past research has found that the experience of state
self-objectification in mixed-gender contexts (stranger and romantic;
Strelan and Pagoudis (2018); Meltzer (2020)) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
for women would suggest no (Calogero, 2004; Gay \& Castano, 2010;
Gervais et al., 2011; Saguy et al., 2010; Yilmaz \& Bozo, 2019), but,
there is evidence that women do objectify other women (Harsey \&
Zurbriggen, 2020; Loughnan et al., 2015a; Puvia \& Vaes, 2013) and it
would be helpful to understand the consequencs of this \emph{intra}group
objectification for women. When women objectify other women, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

There is some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán, Jiao, \& Aubrey,
2020). A recent study has investigated the effect of
self-objectification on the reduction of relationship building skills in
general (including same-sex friendships) (Terán et al., 2020). This link
has also been found in research on stigmatized-stigmatizer interactions
(M. R. Hebl \& Dovidio, 2005; Richeson \& Shelton, 2003; Shelton,
Richeson, \& Salvatore, 2005), and we might view the experience of being
objectified in an interaction as a potential identity threat,
stigmatizing, situation (Miles-McLean et al., 2015; Nadal \& Haynes,
2012). Further, empirical evidence reveals that objectification
manifests through inauthenticity in romantic relationships (Brunell et
al., 2010), an intergroup context at least in heterosexual
relationships. Evidence for otehr adverse consquences of interpersonal
objectification include reductions in career aspirations (Garcia et al.,
2016) and a decrease in concentration and impairment in female cognitive
performance (Kahalon, Shnabel, \& Becker, 2018; D. M. Quinn, Chaudoir,
\& Kallen, 2011). These are also negative conseqences found for women
under steroetype threat (Davies, Spencer, \& Steele, 2005). When a woman
is objectified by a man, and subsequently experiences
self-objectification, the intergroup nature of this encounter might
trigger threat. Thus, perhaps there are fewer negative consequences when
a woman is objectified by another woman. That is, the intergroup threat
literature might predict that woman-woman interpersonal objectification
processes diverge from mixed-gender interpersonal objectification
processes precisely because they are not intergroup interactions (at
least with respect to gender identity).

In summary, a woman objectified by another woman may not be having the
same negative consequences that cascade from situations that trigger
group-based identity threat (Deaux \& Major, 1987; Dovidio, Hebl,
Richeson, \& Shelton, 2006; M. R. Hebl \& Dovidio, 2005); however,
feeling like a body rather than a full human (higher SSO) in any
interaction, whether intergroup or intragroup, may be enough to reduce
women's feelings of authenticity and social competence, regardless of
the gender of the objectifyer (Terán et al., 2020; Tolman, Impett,
Tracy, \& Michael, 2006).

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on intragroup woman objectification is small, the results are
mixed and do not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would be related to state self-objectification (SSO). We also
expected that TSO would moderate this relationship, amplifying the
positive assocaition between being objectified and SSO for women higher
in TSO. Here we would like to note that Puvia and Vaes (2013) would
alternatively predict that women's tendency to self-objectify (TSO)
leads them to objectify other women (more precisely, to dehumanize a
sexualized woman). In addition, Strelan and Hargreaves (2005) and Harsey
and Zurbriggen (2020) would also predict the TSO to
other-objectification link, further theorizing that this relationship is
mediated by the woman's own state self-objectification (SSO). Thus, the
limited but extant literature on woman-woman objectification points to a
possible alternative model from the model tested in the current study.
Where appropriate, we report results considering the causal directions
implied by this alternative mediation model.

We also tested whether SSO would, in turn, lead to feelings of
inauthenticity. The investigation of the SSO to inauthenticity
connection is considered exploratory given the lack of support for this
connection from the literature on identity threat in \emph{intr}agroup
interactions (Rollero, 2016) but support for this connection in the
objectification literature (Garcia et al., 2016; Terán et al., 2020).
Regardless of whether there was an association between SSO and
inauthenticity, we hypothesized that feelings of inauthencity would be
associated with reduced feelings of agency in romantic relationships,
reduced career aspirations, and reduced cognitive performance. In
summary, we expected to find a positive relationship between
other-objectification by one's partner and state self-objectification.
We also expected to find a negative relationship between self-state
objectification and interaction authenticity, and that interaction
authenticity will be positively related to cognitive performance,
relationship agency, and career aspirations.

@@

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Westurn culture and beyond, women need to
look attractive to obtain and maintain successful relationships, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners (i.e., their sexual attractiveness; Puvia \& Vaes,
2013)---indeed, women may be unfortunately quite used to thinking about
their own \enquote{datability,} and we suspect this this habitual
thought pattern will translate to their thoughts about other women.
Analyses comparing levels of other-objectification between heterosexual
and non-heterosexual women are provided in the measures section below.

@@

Thirty-two previously unacquainted dyads of self-identifying women
participated in this study. Data from two demographically similar higher
education institutions in the Northeast United States were combined to
create the final analysis sample (\emph{N =} 64) used in this study. In
the measures section that follows we refer to them as Sample 1 and
Sample 2. Sample 1 (\emph{N =} 24) is from a co-ed liberal arts college
and Sample 2 (\emph{N =} 40) is from a women's liberal arts college.
More specifically, twelve of the pairs were students at a co-ed liberal
arts college, while the remaining twenty pairs attended a women's
liberal arts college.\footnote{Initially, data was collected from both
  same-gender and mixed-gender dyads at both institutions. Sample 1
  originally consisted of 22 pairs, 12 men and 32 women. In Sample 2
  there were 23 pairs made up of 43 women and one man, as well as two
  participants who did not identify as either a woman or man. To
  investigate only same-gender pairs of women, we limited participant
  data to women in both samples.} All results presented below are from
models including sample as a control variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were different racial minority group pairs.
64.06\% of the sample identified as heterosexual, and 25\% identified as
gay, lesbian or bisexual. See the analysis of differences between
heteroseuxal and non-heterosexual women in levels of
other-objectification of their female partners in the measures section
below.

@@

The following measures were collected in the order they are presented
following the interaction. Correlations and descriptive statistics of
all study variables appear in Table~\ref{tab:corrtable}.

@@

Trigrams from the Remote Associates Task (McFarlin \& Blascovich, 1984)
were utilized to assess cognitive performance after the interaction. Ten
items were selected and presented to participants. For example, the
correct answer for the trigram \enquote{Quack: Pond: Waddle} would be
\enquote{Duck}. Participants are limited to 30 seconds to provide their
answers. For every correct answer, 1 point is given. The mean score was
5.03 (SD = 2.29). Cognitive performance was measured first directly
after the intercation in order to measure potential immediate detriments
to performance (Garcia et al., 2016).

@@

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts during the interaction in relation to multiple
characteristics of their partner (Garcia et al., 2016). Questions
included aspects of their partner's internal traits such as personality,
friends, family, and extracurricular interests, as well as external
traits such as body, appearance, clothing, and body parts. All questions
were rated on a scale from 1 (not at all) to 7 (constantly).
Objectification was measured by taking the difference between the
average frequency of thought about their partner's external traits
(\(\alpha\) = 0.79 for Sample 1, \(\alpha\) = 0.79 for Sample 2) and
frequency of thought about their partner's internal traits (\(\alpha\) =
0.79 for Sample 1, \(\alpha\) = 0.76 for Sample 2). A positive score in
this scale would indicate that the participant thought about their
partner's external traits more than the partner's internal traits, and a
negative score would indicate the opposite.

@@

To assess the magnitude to which individuals felt comfortable in the
interaction and perceived the interaction to be authentic, we asked
participants to rate the extent to which they felt comfortable, happy,
friendly, warm, easygoing, sincere, and authentic on a scale ranging
from 1 (not at all) to 7 (very much), much like (Garcia et al., 2016).
Participants were additionally asked to rate their interaction partner's
authenticity as well as their own: \enquote{Do you think your partner
was authentic during your interaction?} and \enquote{Were you authentic
during your interaction?} These questions were ranked on a scale from 1
(not at all) to 7 (very much). These were combined to form the
authenticity scale (\(\alpha\) = 0.91 for Sample 1, \(\alpha\) = 0.91
for Sample 2).

@@

A scale was used from Garcia et al. (2016) to assess how much agency an
individual believes they would possess in future romantic relationships.
Participants were asked how likely it was that they would do the
following: \enquote{ask someone out on a date,} \enquote{open the door
for your date,} \enquote{pay for a date,} \enquote{ask your
boyfriend/girlfriend to marry you,} \enquote{initiate sex with your
girlfriend/boyfriend,} \enquote{initiate condom use during sex,}
\enquote{surprise your boyfriend/ girlfriend with a gift,} and
\enquote{ask your girlfriend/boyfriend to move with you to a new place.}
Responses were measured on a scale ranging from 1 (not at all likely) to
7 (extremely likely). The scale originally had 9 items, but the 9th item
had low correlations with the remaining items, ranging from .02 to .30
for the first sample, and .04 to .30 for the second sample. The item was
intended to be reverse coded, but correlations were still low enough to
make the scale unreliable. Therefore, the ninth item was removed. As a
result, the 8-item scale had moderately high reliability for both
samples (\(\alpha\) = 0.72 for Sample 1, (\(\alpha\) = 0.74 for Sample
2).

@@

We used R (Version 3.5.2; R Core Team, 2017) and the R-packages
\emph{tidyverse} (Version 1.2.1; Wickham, 2017), \emph{psych} (Version
1.8.4; Revelle, 2017), and \emph{nlme} (Version 3.1.137; Pinheiro,
Bates, DebRoy, Sarkar, \& R Core Team, 2017) for our analyses. Further,
we used the R-packages \emph{papaja} (Version 0.1.0.9842; Aust \& Barth,
2018), \emph{apaTables} (Version 2.0.5; Stanley, 2018), with
\emph{knitr} (Version 1.25; Xie, 2015) to create a fully reproducible
manuscript in \emph{rmarkdown} (Version 1.15; Xie, Allaire, \&
Grolemund, 2018).
